A Horrocks-type theorem for even orthogonal $K_2$

### 2020-03-28 dl-paper.tex
Fix 8-9 of 3.5

\begin{df} \label{df:root-subsets}
Recall that a root subset $S \subseteq \Phi$ is called {\it closed} if $\alpha, \beta \in S$ and $\alpha+\beta\in \Phi$ imply $\alpha + \beta \in S$. 
Recall that a closed root subset $S$ is called {\it parabolic} (resp. {\it reductive}, resp. {\it special}) if $S \cup -S = \Phi$ (resp. $S = -S$, resp. $S \cap (-S) = \varnothing$).
The {\it special part} $\Sigma_S$ of a parabolic subset $S$, by definition, consists of all $\alpha \in S$ such that $-\alpha \not\in S$.
\end{df}

@@

\subsection{Computation of the kernel of the map of evaluation at $0$.}
Let $A$ be a local ring with maximal ideal $M$.
The aim of this subsection is to describe a generating set for the kernel of the map $ev_{X=0}^*\colon\overline{\St}(\Phi, A[X], M[X]) \to \overline{\St}(\Phi, A, M)$
induced by the ring homomorphism of evaluation at $0$. We denote this kernel by $K(A[X], M[X])$.

@@

\subsection{The subgroups \texorpdfstring{$P_\alpha(0)$}{Pa(0)}, \texorpdfstring{$P_\alpha(*)$}{Pa(*)} and their properties.} \label{sec:Pa0-basic}
Let $\Phi$ be a root system and $\alpha$ be an element of $\Phi$.
Consider the following subsets of $\Phi$:
\begin{align} Z_+(\alpha) & = \{ \beta \in \Phi \mid \langle \alpha, \beta \rangle > 0 \}, \\
   Z_0(\alpha) & = \{ \beta \in \Phi \mid \alpha + \beta \not\in \Phi,\ \langle \alpha, \beta \rangle = 0 \}, \\
   Z(\alpha)   & = Z_0(\alpha) \sqcup Z_+(\alpha). \end{align}
Clearly, $Z_0(\alpha)$ (resp. $Z_+(\alpha)$) is a reductive (resp. special) subset of $\Phi$ (see~\cref{df:root-subsets} for the terminology).

@@

\begin{df} \label{defP0}
Let $M$ be an ideal of a local ring $A$. We denote by $P_{\alpha, M}(0)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by the images of the following five types of elements of $\overline{\St}(\Phi, A[X], M[X])$ under the map $j_+ \colon \St(\Phi, A[X]) \to \St(\Phi, A[X, X^{-1}])$:
\begin{enumerate} \item $z_{\beta}(Xf, \xi),\ \beta \in \Phi \text{ such that }\alpha + \beta \in \Phi\text{ or } \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, X\xi),\ \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, \xi),\ \beta \perp \alpha;$
 \item $x_{-\alpha}(X^2f);$
 \item $x_{\alpha}(f)$. \end{enumerate}
In the above formulae and for the rest of this subsection $f \in M[X]$ and $\xi \in A[X]$. Notice also that by~\eqref{Zdef} one has $z_\beta(f, 0) = x_\beta(f)$.

We also denote by $P_{\alpha, M}(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_{\alpha, M}(0)$ and the elements $x_{-\alpha}(mX)$, $m \in M$.
\end{df}

@@

\begin{rem}\label{rem:c-DS} Notice that $P_{\alpha, M}(0)$ contains the elements $c_{\beta}(f, X\xi)$ for all $\beta \in Z(\alpha)$ (they can be factored as products of two elements of type 2 or 3). It is also easy to check that $P_{\alpha, M}(0)$ contains the image of $Z_\alpha(A, M)$ under the natural embedding of $\St(\Phi, A, M) \hookrightarrow \St(\Phi, A[X, X\inv])$.
In particular, if $M$ is the maximal ideal of $A$, the subgroup $P_{\alpha, M}(0)$ contains relative Dennis--Stein and Steinberg symbols. \end{rem}

The following lemma shows that $P_{\alpha, M}(*)$ is sufficiently large.
\begin{lemma} \label{Pstar-large} Suppose that $(A, M)$ is a local pair. Then the subgroup $P_{\alpha, M}(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof} Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$ hence by~\cref{thm:Stepanov} $P_\alpha(*)$ contains whole $j_+(\overline{\St}(\Phi, A[X], XM[X]))$. The required assertion now follows from~\cref{Kgen-strong} and the preceding remark. \end{proof}

>>>>

\begin{df} \label{df:root-subsets}
Recall from~\cite[Ch.~VI, \S~1.7]{Bou81} that a root subset $S \subseteq \Phi$ is called {\it closed} if $\alpha, \beta \in S$ and $\alpha+\beta\in \Phi$ imply $\alpha + \beta \in S$. 
Recall that a closed root subset $S$ is called {\it parabolic} (resp. {\it symmetric}, resp. {\it special}) if $S \cup -S = \Phi$ (resp. $S = -S$, resp. $S \cap (-S) = \varnothing$).
The {\it special part} $\Sigma_S$ of a parabolic subset $S$, by definition, consists of all $\alpha \in S$ such that $-\alpha \not\in S$.
\end{df}

@@

\subsection{Computation of the kernel of the map of evaluation at $0$.}\label{sec:computationOfK}
Let $A$ be a local ring with maximal ideal $M$.
The aim of this subsection is to describe a generating set for the kernel of the map $ev_{X=0}^*\colon\overline{\St}(\Phi, A[X], M[X]) \to \overline{\St}(\Phi, A, M)$
induced by the ring homomorphism of evaluation at $0$. We denote this kernel by $K(A[X], M[X])$.

@@

\subsection{The subgroups \texorpdfstring{$P_\alpha(0)$}{Pa(0)}, \texorpdfstring{$P_\alpha(*)$}{Pa(*)} and their properties.} \label{sec:Pa0-basic}
Let $\Phi$ be a root system and $\alpha$ be an element of $\Phi$.
Consider the following subsets of $\Phi$:
\begin{align} Z_+(\alpha) & = \{ \beta \in \Phi \mid \langle \alpha, \beta \rangle > 0 \}, \\
   Z_0(\alpha) & = \{ \beta \in \Phi \mid \alpha + \beta \not\in \Phi,\ \langle \alpha, \beta \rangle = 0 \}, \\
   Z(\alpha)   & = Z_0(\alpha) \sqcup Z_+(\alpha). \end{align}
Clearly, $Z_0(\alpha)$ (resp. $Z_+(\alpha)$) is a symmetric (resp. special) subset of $\Phi$ (see~\cref{df:root-subsets} for the terminology).

@@

\begin{df} \label{defP0}
Let $M$ be an ideal of a local ring $A$. We denote by $P_{\alpha, M}(0)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by the images of the following five families of elements of $\overline{\St}(\Phi, A[X], M[X])$ under the map $j_+ \colon \St(\Phi, A[X]) \to \St(\Phi, A[X, X^{-1}])$:
\begin{enumerate} 
 \item $z_{\beta}(Xf, \xi),\ \beta \in \Phi \text{ such that }\alpha + \beta \in \Phi\text{ or } \alpha - \beta \in \Phi,\ f \in M[X],\ \xi \in A[X];$
 \item $z_{\beta}(f, X\xi),\ \alpha - \beta \in \Phi,\ f\in M[X],\ \xi\in A[X];$
 \item $z_{\beta}(f, \xi),\ \beta \perp \alpha,\ f\in M[X],\ \xi\in A[X];$
 \item $x_{-\alpha}(X^2f),\ f \in M[X];$
 \item $x_{\alpha}(f),\ f \in M[X]$. \end{enumerate}
Notice that in the above definition we use letter $f$ (resp. $\xi$) to denote an element of $M[X]$ (resp. $A[X]$).
We adhere to this convention until the end of this section.
Notice also that by~\eqref{Zdef} one has $z_\beta(f, 0) = x_\beta(f)$.

We denote by $P_{\alpha, M}(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_{\alpha, M}(0)$ and the elements $x_{-\alpha}(mX)$, $m \in M$.
\end{df}

@@

\begin{rem}\label{rem:c-DS} By~\eqref{Cdef} one has $c_\beta(f, X\xi) = x_{\beta} (f) \cdot z_{\beta}(-f, -X\xi)$. This implies that for $\beta \in Z_+(\alpha)$ (resp. $\beta \in Z_0(\alpha)$) the element $c_\beta(f, X\xi)$ is a product of two generators of type 2 (resp. 3).
Thus, the elements $c_{\beta}(f, X\xi)$ lie in $P_{\alpha, M}(0)$ for all $\beta \in Z(\alpha)$.

It is also easy to check that $P_{\alpha, M}(0)$ contains the image of $Z_\alpha(A, M)$ under the natural embedding of $\overline{\St}(\Phi, A, M) \hookrightarrow \St(\Phi, A[X, X\inv])$.
In particular, if $M$ is the maximal ideal of $A$, the subgroup $P_{\alpha, M}(0)$ contains relative Dennis--Stein and Steinberg symbols. \end{rem}

The following lemma shows that $P_{\alpha, M}(*)$ is sufficiently large.
\begin{lemma} \label{Pstar-large} Suppose that $(A, M)$ is a local pair. 
Then the subgroup $P_{\alpha, M}(*)$ contains the image in $\St(\Phi, A[X, X^{-1}])$ of the subgroup $K(A[X], M[X]) \leq \overline{\St}(\Phi, A[X], M[X])$ defined in the beginning of~\cref{sec:computationOfK}. \end{lemma}
\begin{proof} Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$, therefore by~\cref{thm:Stepanov} $P_\alpha(*)$ contains all of $j_+(\overline{\St}(\Phi, A[X], XM[X]))$. The required assertion now follows from~\cref{Kgen-strong} and~\cref{rem:c-DS}. \end{proof}
